An event-sourced and distributed database with IPLD and materialized views

### 2019-09-19 draft.tex
links a bit more; fixes a few things
Signed-off-by: andrew <andrew@textile.io>

\begin{definition} (Logs or Append-only log). A log is a registry of database transactions that is read sequentially (normally ordered by time) from beginning to end. In distributed systems, logs are often treated as append-only, where changes or updates can only be added to the set and never removed.  \end{definition}

@@

A key principal of ES and append only logs is that all changes to application state are stored as a sequence of events. Because any given state is simply the result of a series of atomic updates, the log can be used to reconstruct past states or process retroactive updates [ref]. The same principal means a log can be viewed as a mechanism to support an infinite number of valid state interpretations (see section  \ref{sec:viewsprojections}). In other words, with minimal conformity, a single log can model multiple application states [ref].

\subsubsection{Views \& Projections} \label{sec:viewsprojections}

@@

\subsubsection{Eventual Consistency}

@@

\begin{definition} (CRDT). A conflict-free replicated data type (CRDT) assures eventual consistency through optimistic replication (i.e. all new updates are allowed) and eventual merging. CRDTs rely on data structures that are mathematically guaranteed to resolve concurrent updates the same way regardless of the order those events were received. \end{definition} \label{def:crdt}

How a system provides eventual consistency is often decided based on the intended use of the system. Two well-documented categories of solutions include logs (see Definition X.x.x), sequences of deterministic changes, and CRDTs (see Definition XXX). Logs work best when there is only a single-writer or complete event causality is known. In many distributed systems with multiple log writers, a minimum requirement for synchronization through logs is that the essential order of events is respected and can be determined [3, 8]. For these cases, logical clocks are a useful tool for eventual consistency and total ordering [ref]. However, some scenarios (e.g., temporarily missing events, or ambiguous order) can force a replica into a state that cannot be later resolved without costly recalculation. In specific cases, CRDTs can provide an alternative to log-based consensus (see below).

@@

CRDTs (see \ref{def:crdt}) are one way to achieve strong eventual consistency, where once all replicas have received the same events, they will arrive at the same final state, regardless of ordering. A review of the types of possible CRDTs is beyond the scope of this paper, however, it is important to note their role in eventually consistent systems and how they relate to clock-based event ordering. See for example [Merkle Clock ref] and [ref] for informative reviews of these types of data structures.

@@

In the figure above, two replicas (top row, bottom row) are attempting to write (left to right) events to the same Merkle-Clock. After the first replica writes event A, the second writes event A’ and properly links to A. At that point, the two replicas stop receiving events from one another. To a third replica that does continue to receive events, there would now be two independent heads, 1 and 1’. For the third replica, resolving these two logs of events may be costly (many updates happened since the last common node) or impossible (parts of the chain may not be available on the network). See section 2.2.3. Eventual Consistency.

@@

Event-driven systems (e.g., event sourcing) often make use of ACLs with some distinct properties. The ACL of an event-driven system is usually a list of access rules built from a series of events. For example, the two events, "grant Bob write access" and "revoke read access from Alice" would together result in a final ACL state where, Bob has read and write access, but Alice does not (see Views \& Projections in Data Synchronization).

@@

\begin{definition}
(Reader). Any Peer capable of reading a Log. Practically speaking, this means any Peer with the Log’s Read key (see section 3.1.3 Keys \& Encryption).
\end{definition}

@@

Together with a cryptographic signature, an Event is written to a log with an additional node (see Figure X.X.) enabling log verification by readers (see section 3.1.3. Keys \& Encryption).

@@

Unlike peer multiaddresses, Log addresses are not stored in the global IPFS distributed hash table [ref] (DHT). Instead, they are collected from Log Events (see section 3.3 Use Cases). This is in contrast to mutable data via IPNS for example, which requires querying the network (DHT) for updates. Instead, updates are requested directly from the (presumably trusted) peers that produced them, resulting in a hybrid of content-addressed Events arranged over a data-feed like topology. Log addresses are recorded in an address book, similar to IPFS Peer address book:

@@

\begin{definition}Follower). Log Writers can designate other IPFS Peers to “follow” a Log, potentially replicating and/or republishing Events. A Follower is capable of receiving Log updates and traversing linkages via the Follow Key (see section 3.1.3 Keys \& Encryption), but is not able to read the Log’s contents. Followers should be server-based, i.e., always online and behind a public IP address.\end{definition}

@@

In practice, Writers are solely responsible for announcing their Log’s addresses. This ensures a conflict-free address list without additional complexity. Some Followers may be in the business of replicating Logs (see section 5.1. Textile: The Thread \& Bot Network), in which case Writers will announce the additional Log address to Readers. This allows them to pull (or subscribe to push-based) Events from the Follower’s Log address when the Writer is offline or unreachable.

@@

\subsubsection{Keys \& Encryption}

Textile Logs are designed to be shared, composed, and layered into datasets. Therefore, Logs are encrypted by default in a manner that enables access control (see section 3.4. Access Control) and the Follower mechanism discussed in the previous section.

@@

\begin{definition}
(Access-Controlled): This variant declares that consumers should assume an access control list is composable from Log Events. The ACL represents a permissions rule set that must be applied when reading data (see section 3.4.2. Access Control).
\end{definition}

@@

\subsubsection{Log Synchronization}

@@

\begin{definition}
(Invite). An Event containing a mapping of Log IDs to Key Sets, which can be used to join a Thread. Threads backed by an access control list (see section 3.4.2. Access Control) will also include the current ACL for the Thread in an Invite. This enables Peers to invite others to only read or follow a Thread, instead of becoming a full-on collaborator, i.e., a new Log Writer.
\end{definition}

@@

Once a Peer has accepted an Invite, it will receive new Events from Log Writers. In cases where the invitee becomes a collaborator, i.e., a Writer, it is also responsible for sending its own Events out to the network.

Sending

@@

\begin{enumerate}
\item \label{Perf1}New Events may be pushed to web-based participants over a WebSocket.
\item \label{Perf2}New Events may be pushed to the Thread’s other Log Writers via federated notification services like Apple Push Notification Service (APNS) [ref], Google Cloud Messaging (GCM) [ref], Firebase Cloud Messaging (FCM) [ref], and/or Windows Notification Service (WNS) [ref].
\item \label{Perf3} New Events may trigger web-hooks [ref] enabling complex workflows (see Section X below).
\end{enumerate}

Receiving

@@

The notion of the Follow key (see section 3.1.3 Keys \& Encryption) makes duplicating all Log Events trivial. This allows any Peer in the network to be granted the responsibility of replicating data from another Peer without having read access to the data contained within the Log entries. This type of Log replication can act as a data backup mechanism. It can also be used to build services that react to Log Events, potentially pushing data to disparate, non-Textile systems, especially if the replication service is granted read access to the Log Events (see section 3.2.3 Log Synchronization).

@@

\subsubsection{Access Control}

@@

\bibliographystyle{compj}
% \bibliography{ModellingBidders}
\input{sample.bbl}

>>>>

\begin{definition} (Logs or Append-only log). A log is a registry of database transactions that is read sequentially (normally ordered by time) from beginning to end. In distributed systems, logs are often treated as append-only, where changes or updates can only be added to the set and never removed.  \end{definition} \label{def:Logs}

@@

A key principal of ES and append only logs is that all changes to application state are stored as a sequence of events. Because any given state is simply the result of a series of atomic updates, the log can be used to reconstruct past states or process retroactive updates [ref]. The same principal means a log can be viewed as a mechanism to support an infinite number of valid state interpretations (see section  \ref{sec:ViewsProjections}). In other words, with minimal conformity, a single log can model multiple application states [ref].

\subsubsection{Views \& Projections} \label{sec:ViewsProjections}

@@

\subsubsection{Eventual Consistency} \label{sec:EventualConsistency}

@@

\begin{definition} (CRDT). A conflict-free replicated data type (CRDT) assures eventual consistency through optimistic replication (i.e. all new updates are allowed) and eventual merging. CRDTs rely on data structures that are mathematically guaranteed to resolve concurrent updates the same way regardless of the order those events were received. \end{definition} \label{def:CRDT}

How a system provides eventual consistency is often decided based on the intended use of the system. Two well-documented categories of solutions include logs (see Definition  \ref{def:Logs}), sequences of deterministic changes, and CRDTs (see Definition section  \ref{def:CRDT}). Logs work best when there is only a single-writer or complete event causality is known. In many distributed systems with multiple log writers, a minimum requirement for synchronization through logs is that the essential order of events is respected and can be determined [3, 8]. For these cases, logical clocks are a useful tool for eventual consistency and total ordering [ref]. However, some scenarios (e.g., temporarily missing events, or ambiguous order) can force a replica into a state that cannot be later resolved without costly recalculation. In specific cases, CRDTs can provide an alternative to log-based consensus (see below).

@@

CRDTs (see Definition \ref{def:CRDT}) are one way to achieve strong eventual consistency, where once all replicas have received the same events, they will arrive at the same final state, regardless of ordering. A review of the types of possible CRDTs is beyond the scope of this paper, however, it is important to note their role in eventually consistent systems and how they relate to clock-based event ordering. See for example [Merkle Clock ref] and [ref] for informative reviews of these types of data structures.

@@

In the figure above, two replicas (top row, bottom row) are attempting to write (left to right) events to the same Merkle-Clock. After the first replica writes event A, the second writes event A’ and properly links to A. At that point, the two replicas stop receiving events from one another. To a third replica that does continue to receive events, there would now be two independent heads, 1 and 1’. For the third replica, resolving these two logs of events may be costly (many updates happened since the last common node) or impossible (parts of the chain may not be available on the network). See Section  \ref{sec:EventualConsistency}.

@@

Event-driven systems (e.g., event sourcing) often make use of ACLs with some distinct properties. The ACL of an event-driven system is usually a list of access rules built from a series of events. For example, the two events, "grant Bob write access" and "revoke read access from Alice" would together result in a final ACL state where, Bob has read and write access, but Alice does not (see Section  \ref{sec:ViewsProjections}).

@@

\begin{definition}
(Reader). Any Peer capable of reading a Log. Practically speaking, this means any Peer with the Log’s Read key (see Section  \ref{sec:KeysEncryption}).
\end{definition}

@@

Together with a cryptographic signature, an Event is written to a log with an additional node (see Figure X.X.) enabling log verification by readers (see Section  \ref{sec:KeysEncryption}).

@@

Unlike peer multiaddresses, Log addresses are not stored in the global IPFS distributed hash table [ref] (DHT). Instead, they are collected from Log Events (see Section  \ref{sec:UseCases}). This is in contrast to mutable data via IPNS for example, which requires querying the network (DHT) for updates. Instead, updates are requested directly from the (presumably trusted) peers that produced them, resulting in a hybrid of content-addressed Events arranged over a data-feed like topology. Log addresses are recorded in an address book, similar to IPFS Peer address book:

@@

\begin{definition}Follower). Log Writers can designate other IPFS Peers to “follow” a Log, potentially replicating and/or republishing Events. A Follower is capable of receiving Log updates and traversing linkages via the Follow Key (see Section  \ref{sec:KeysEncryption}), but is not able to read the Log’s contents. Followers should be server-based, i.e., always online and behind a public IP address.\end{definition}

@@

In practice, Writers are solely responsible for announcing their Log’s addresses. This ensures a conflict-free address list without additional complexity. Some Followers may be in the business of replicating Logs (see section  \ref{sec:Bots}), in which case Writers will announce the additional Log address to Readers. This allows them to pull (or subscribe to push-based) Events from the Follower’s Log address when the Writer is offline or unreachable.

@@

\subsubsection{Keys \& Encryption} \label{sec:KeysEncryption}

Textile Logs are designed to be shared, composed, and layered into datasets. Therefore, Logs are encrypted by default in a manner that enables access control (see Section  \ref{sec:AccessControl}) and the Follower mechanism discussed in the previous section.

@@

\begin{definition}
(Access-Controlled): This variant declares that consumers should assume an access control list is composable from Log Events. The ACL represents a permissions rule set that must be applied when reading data (see Section  \ref{sec:AccessControl}).
\end{definition}

@@

\subsubsection{Log Synchronization} \label{sec:LogSync}

@@

\begin{definition}
(Invite). An Event containing a mapping of Log IDs to Key Sets, which can be used to join a Thread. Threads backed by an access control list (see Section  \ref{sec:AccessControl}) will also include the current ACL for the Thread in an Invite. This enables Peers to invite others to only read or follow a Thread, instead of becoming a full-on collaborator, i.e., a new Log Writer.
\end{definition}

@@

Once a Peer has accepted an Invite, it will receive new Events from Log Writers. In cases where the invitee becomes a collaborator, i.e., a Writer, it is also responsible for sending its own Events out to the network.\\

\textbf{Sending} \\

@@

\begin{enumerate}
\item \label{Perf1}New Events may be pushed to web-based participants over a WebSocket.
\item \label{Perf2}New Events may be pushed to the Thread’s other Log Writers via federated notification services like Apple Push Notification Service (APNS) [ref], Google Cloud Messaging (GCM) [ref], Firebase Cloud Messaging (FCM) [ref], and/or Windows Notification Service (WNS) [ref].
\item \label{Perf3} New Events may trigger web-hooks [ref] enabling complex workflows (see Section X below).\\
\end{enumerate}

\textbf{Receiving} \\

@@

The notion of the Follow key (see Section  \ref{sec:KeysEncryption}) makes duplicating all Log Events trivial. This allows any Peer in the network to be granted the responsibility of replicating data from another Peer without having read access to the data contained within the Log entries. This type of Log replication can act as a data backup mechanism. It can also be used to build services that react to Log Events, potentially pushing data to disparate, non-Textile systems, especially if the replication service is granted read access to the Log Events (see section  \ref{sec:LogSync}).

@@

\subsubsection{Access Control} \label{sec:AccessControl}

@@

\bibliographystyle{compj}
% \bibliography{ModellingBidders}
\input{draft.bbl}
